---
title: "Owner Spend Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

In this workbook we'll build a model for owner spend in the month of 
August 2015. (The Wedge did a remodel project in February of 2016, 
so we want to be careful using time after that for our analysis.)

Let's start by reading in a file of spend by owner and department by 
month and year. Feel free to substitute your file in for the 
one I'm providing--you should have something like this built for
part 2 of the Wedge project.


```{r input, cache=TRUE, echo=FALSE, include=FALSE}
# Caching keeps this from running every time you Knit.
d <- data.table::fread("20161107_owner_month_dept.txt",header=T)
```

```{r head, results="asis",echo=FALSE}
knitr::kable(head(d),caption="Head of the Owner Data")

```

You should be pretty familiar with those fields except for "Segment". 
That field is built based on some rules that we built back in 2015
to split owners into various categories. Let me know if you'd like to hear
more about the process. 

As always, let's do some descriptive statistics. Here's what
`describe` has to say for us. We're going to cut down the
data to `segment != 'light' & year == 2015` just to make
the code run snappier.

``` {r desc, echo=F, warning=F}
d <- subset(d,segment != 'light' & year == 2015)
Hmisc::describe(d)
```

There are some *very* large spends in there. So that our regression
model isn't unduly influenced by these non-individual owners, 
let's drop anyone with a department-level spend of more than $1,000
in a month. This isn't perfect, and if we worked for the Wedge there
would be better ways to clean this data, but it will get us started.
``` {r echo=F}
questionable_spenders <- unique(d$owner[d$spend > 1000])

d <- d[!(d$owner %in% questionable_spenders),]
```


## Building Regression Data
We'd like to build a regression model that has
spend in August 2015 as the response variable. Our
explantory variables will be:
* Spend in July
* Spend in June
* Number of departments shopped in previous seven months
* Indicator variables for July shopping in Produce, 
Packaged Grocery, and Supplements. 

There are a lot of ways to build up this data set, but since
we're easing our way back into R, I'll do it in a slow-but-clear
way so we can see what's happening. Once we get `dplyr` (next week)
we'll see some faster ways.

``` {r data_build, cache=T}
reg.d <- data.frame(owner=unique(d$owner[d$month==8 & d$year==2015]), # get owners
                    august=0, # placeholders
                    july=0, 
                    june=0,
                    depts=0,
                    produce=0,
                    pck_groc=0,
                    supplements=0)

for (i in 1:nrow(reg.d)) {
  this.owner <- reg.d$owner[i]
  this.data <- d[d$owner==this.owner,]
  
  reg.d$august[i] <- sum(this.data$spend[this.data$month==8])
  reg.d$july[i] <- sum(this.data$spend[this.data$month==7])
  reg.d$june[i] <- sum(this.data$spend[this.data$month==6])
  reg.d$depts[i] <- length(unique(this.data$dept_no[this.data$month < 8]))
  reg.d$produce[i] <- ifelse("PRODUCE" %in% unique(this.data$department[this.data$month==7]),1,0)

  reg.d$pck_groc[i] <- ifelse("PACKAGED GROCERY" %in% unique(this.data$department[this.data$month==7]),1,0)

  reg.d$supplements[i] <- ifelse("SUPPLEMENTS" %in% unique(this.data$department[this.data$month==7]),1,0)

}

```

## Plotting the data
It's always a good idea to look at the data before we start 
doing some modeling. Let's look at the various columns of `reg.d` 
versus the response variable, `august`. 

```{r plots, echo=FALSE}
plot(reg.d$july,reg.d$august,xlab="july",ylab="august")
plot(reg.d$june,reg.d$august,xlab="june",ylab="august")

boxplot(reg.d$august ~ reg.d$depts,xlab="depts",ylab="august")

boxplot(reg.d$august ~ reg.d$produce,xlab="produce",ylab="august")
boxplot(reg.d$august ~ reg.d$pck_groc,xlab="packaged grocery",ylab="august")
boxplot(reg.d$august ~ reg.d$supplements,xlab="supplements",ylab="august")

```

In the future when we make charts like this we can make them look
much better (particularly using `ggplot`). We haven't covered that
yet, though, so I'll just make one for you to show you how I'd 
prepare these in a more publication-worthy way.
```{r gg_plot}
library(ggplot2)
library(scales)

p <- ggplot(reg.d,
            aes(x=july,y=august)) + 
  geom_point(alpha=0.2) + 
  labs(x="July Spend",y="August Spend") +
  theme_bw() +
  scale_x_continuous(label=dollar) + 
  scale_y_continuous(label=dollar) + 
  stat_smooth(method="lm",col="red") + 
  stat_smooth(col="blue",se=F)

print(p)

```

## Building a Model
Building models in R are pretty easy. Let's create one and look
at the summary-level information

``` {r model_1}
lm.1 <- lm(august ~ july + june + depts + 
             produce + pck_groc + supplements,
           data=reg.d)

summary(lm.1)
```

The $R^2=0.7$, which is pretty respectable and our residual 
standard error is $122, which, again, seems pretty solid until 
we remember that the mean August spend is $193, so we're missing
by a pretty decent amount. 

Spend some time thinking about what these coefficients mean. What
would the estimate be for someone who spent $100 in June and July, 
shopped 10 departments, and shopped PRODUCE and SUPPLEMENTS in July?
You can write some R code here to answer the question.

## Analyzing Residuals
Let's add the residuals of the model to our regression data. 
``` {r add_resid} 
reg.d$resids <- residuals(lm.1)
Hmisc::describe(reg.d$resids)

knitr::kable(subset(reg.d,resids < -500))
knitr::kable(subset(reg.d,resids > 800))

```

How would you describe what we're seeing in this large and small 
residuals? What could you do to try to improve our model fit?

## Future Work
Here are some ways to extend this work if we've got more time:
* Whatever you came up with immediately above! :-)
* Extending to a rolling time horizon rather than having August 2015
as the only response variable.
* Exploring additional covariates. 